\chapter{Background}

\section{Executing Code on the Browser}

The only PL that standard web browsers (\emph{Google Chrome, Mozilla Firefox, Safari, Edge, Internet Explorer, etc.}) can execute is JavaScript. There is a movement to bring other PLs to the browser via a technology called WebAssembly, but this has not fully come to fruition yet.  
In addition to this restriction on PLs, each browser instance limits executing code to a single, non-blocking thread.

\subsection{Task Queue APIs}\label{background:task-queues}
To overcome the non-blocking limitation of the execution thread the JavaScript environment exposes an API to a simple ''task queue``.
The task queue is exactly what it sounds like, it's a queue of operations that the JavaScript environment maintains and performs them in ``first in first out'' order (roughly). This task queue contains lots of important tasks, such as mouse and keyboard event processing, scrolling, rendering, button animations etc.

The JavaScript runtime allows users to add their own tasks to the queue by passing in \textbf{callbacks} to certain built-in functions. Callbacks are added to the task queue and will be executed at some later stage.
When the JavaScript runtime completes a task, it moves to the next task in the queue.

By design JavaScript does not allow any blocking operations, e.g., \verb|sleep|, or synchronous network calls or IO reads. The task queue API is therefore used to work around this.
Figure \ref{blockingcode} contains some simple code that would work if the browser thread allowed blocking.

\begin{figure}
    \caption{Code if thread can be blocked.}
    \label{blockingcode}
    \begin{lstlisting}[language=javascript]
// Would work in an ideal world
function sleepFor100ms() {
    print('Starting sleep')
    sleep(100)
    print('After sleep')
}
    \end{lstlisting}
\end{figure}

Ideally, this would print ``Starting Sleep'', wait 100ms, then print ``After Sleep''. However, sleeping would be a blocking action on the thread, so this is not allowed. In JavaScript, we must use the \texttt{setTimeout} function, which takes a callback function, and a time to wait before executing it. Figure \ref{setTimeoutCode} shows some code that uses the \texttt{setTimeout} API to add a callback to the task queue.

\begin{figure}
    \caption{Using \texttt{setTimeout}}
    \label{setTimeoutCode}
    \begin{lstlisting}[language=javascript]
// The real way to achieve this
function sleepFor100ms() {
    print('Starting sleep')
    setTimeout(() => print('After sleep'), 100)
}
    \end{lstlisting}
\end{figure}

Similarly, JavaScript uses an API to overcome the single-thread limitation. The developer is given access to certain functions that can use an extra thread for specific tasks, e.g., Fetching web content from a different website. These APIs also use callback functions to specify what should be done with the result of the functions, they perform their requested action and then place the callback function on the task queue.

This reliance on callback functions leads to something that the JavaScript community refers to as \textbf{Callback Hell}\cite{callbackhell}.

\subsection{async / await and Promises}\label{background:asyncawait}

In an attempt to overcome callback hell, recent versions of JavaScript have introduced a concept known as ``async / await''.\cite{es8spec} This feature introduces what is effectively a syntactic sugar over a concept introduced previously, known as \textbf{Promises}.

Promises are a reference to a value that has not yet been computed, but will at some point be computed (resolved). These references are objects that can be manipulated just like any other object.

Mozilla describes Promises as
\begin{quote}
A Promise is a proxy for a value not necessarily known when the promise is created. It allows you to associate handlers with an asynchronous action's eventual success value or failure reason. This lets asynchronous methods return values like synchronous methods: instead of immediately returning the final value, the asynchronous method returns a promise to supply the value at some point in the future.
    \cite{mozillapromises}
\end{quote}

Promises allow us to use an almost monadic programming style to write our asynchronous programs. However, in a similar way to Haskell's ``do'' notation, the async / await syntax allows us to write our asynchronous programs in a more imperative style. Figure \ref{background:figure:async-await} shows an example of using Promises directly vs using the async / await syntactic sugar to write the same function.

\begin{figure}
    \caption{Promises vs async / await}
    \label{background:figure:async-await}
    \begin{minipage}[t]{0.45\textwidth}
        \begin{lstlisting}[language=javascript]
// Async addition using Promises
function doAsyncAddition() {
    return getA().then(a => {
        return getB().then(b => {
            return a + b;
        });
    });
}
        \end{lstlisting}
    \end{minipage}
    \begin{minipage}[t]{0.45\textwidth}
        \begin{lstlisting}[language=javascript]
// Async addition using async / await syntactic sugar.
async function doAsyncAddition() {
    const a = await getA();
    const b = await getB();
    return a + b;
}
        \end{lstlisting}
    \end{minipage}
\end{figure}

\section{TypeScript}
TypeScript is an open-source PL being developed by Microsoft, TypeScript is the primary PL that is used in the implementation of this project. Microsoft describes it as

\begin{quote}
TypeScript is a typed superset of JavaScript that compiles to plain JavaScript.\cite{microsoftts}
\end{quote}

TypeScript allows developers to write JavaScript programs with all the benefits of compile-time static type checking. As TypeScript is compiled to JavaScript it allows the user to use features of JavaScript that might not be widely supported yet, as it will compile them down to simpler features. Unfortunately, as TypeScript is still compiled to JavaScript, we are left with the same limitations as mentioned before.

\section{PEG and CFG Grammars}

Formal grammars are a model for specifying a certain set of strings. Noam Chomsky first formally researched these grammars in 1956\cite{chomskypaper} and shaped how PL syntax would be specified for years to come. PL designers use formal grammars to specify the syntax structure for their PL.

Context-free grammars (CFGs) are the family of grammars that have seen the most use in PL specification. However, recently there has been an uptick in the use of a similar grammar specification known as Parsing Expression Grammars (PEGs). These grammars are very similar to CFGs but are defined in such a way that they cannot be ambiguous.
For the purposes of this project, PEGs provide a powerful way of formalising the syntax of a language. In fact, it is conjectured that PEGs are more powerful than CFGs, but this remains unproven\cite{pegconjecture}.

\section{Parser Generators}

A parser generator is a program that, given a description of some formal grammar, outputs a parser for that grammar. Parser generators have existed since the earliest days of computer programming. One of the earliest major parser generators, YACC, was released in 1975 \cite{yacc}, and was based on LR parsing.

Several parser generators are available for JavaScript, including the excellent \emph{pegjs} parser generator. However, no such parser generator was available for TypeScript, this is what prompted the creation of \tsPEG{}.

\section{VSO vs SVO languages}

In linguistics, there is a concept known as word order, which is studying the order of how the different categories of words appear in sentences. One order that is studied is called the ``constituent word order''. The constituent word order is the order that a verb (V), object (O), and subject (S) appear in a sentence\cite{wordorder}. English is an SVO language, meaning that the subject comes first, then the verb, then the object. For example
\begin{quote}
    The man drove his car
\end{quote}
In this sentence ``the man'' is the subject, ``drove'' is the verb, and ``his car'' is the object, and they appear in the order SVO.

Irish is a VSO language, meaning the verb comes first, then the subject, then the object. The same above sentence in Irish would be
\begin{quote}
    Thiomáin an fear a charr
\end{quote}
In this case, ``Thiomáin'' (drove) is the verb, ``an fear'' (the man) is the subject, and ``a charr'' (his car) is the object.

