\begin{thebibliography}{9}
    \bibitem{csoirish}{
            CSO, \emph{The Irish language}, 2017,
            \url{https://cso.ie/en/media/csoie/releasespublications/documents/population/2017/7._The_Irish_language.pdf}
        }
    \bibitem{languagecolour}{
            \emph{Language regions of brain are operative in color perception}, Wai Ting Siok, Paul Kay, William S. Y. Wang, Alice H. D. Chan, Lin Chen, Kang-Kwong Luke, Li Hai Tan,
            Proceedings of the National Academy of Sciences May 2009, 106 (20) 8140-8145; DOI: 10.1073/pnas.0903627106
        }
    \bibitem{wikipllist}{
            Non-English based programming languages, Wikipedia
            [\url{https://en.wikipedia.org/wiki/Non-English-based_programming_languages}, Accessed on 2020/02/10]
        }
    \bibitem{graphiclearning}{
            Graphic Organisers: A review of Scientifically Based Research, The Insitute for the Advancement of Research in Education at AEL
            [\url{https://www.inspiration.com/sites/default/files/documents/Detailed-Summary.pdf}, Accessed on 2020/02/10]
        }
    \bibitem{yacc}{Johnson, Stephen C. (1975). Yacc: Yet Another Compiler-Compiler}
    \bibitem{microsoftts}{TypeScript, Microsoft, [\url{https://www.typescriptlang.org/}, Accessed on 2020/02/13]}
    \bibitem{es8spec}{ECMAScript® 2017 Language Specification (ECMA-262, 8th edition, June 2017),
        [\url{https://www.ecma-international.org/ecma-262/8.0/index.html}, Accessed 2020/02/10]}
    \bibitem{mozillapromises}{Mozzila Developer Network Documentation, Promises,
        [\url{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Reference/Global_Objects/Promise}, Accessed on 2020/04/13]}
    \bibitem{chomskypaper}{Chomsky, Noam (Sep 1956). ``Three models for the description of language''. IRE Transactions on Information Theory. 2 (3): 113–124. doi:10.1109/TIT.1956.1056813.}
    \bibitem{pegconjecture}{Ford, Bryan p (2004). ``Parsing Expression Grammars: A Recognition Based Syntactic Foundation''. Proceedings of the 31st ACM SIGPLAN-SIGACT Symposium on Principles of Programming Languages. ACM. doi:10.1145/964001.964011. ISBN 1-58113-729-X.}
    \bibitem{wordorder}{Song, Jae Jung (2012), Word Order. Cambridge: Cambridge University Press. ISBN 978-0-521-87214-0 \& ISBN 978-0-521-69312-}
    \bibitem{chinesepl}{World's First Classical Chinese Programming Language, Charles Q. Choi, IEEE Spectrum, [\url{https://spectrum.ieee.org/tech-talk/computing/software/classical-chinese}, Accessed on 2020/03/02]}
    \bibitem{pegerrors}{Rüfenacht, Michael, Oscar Nierstrasz, and Jan Kurš. "Error handling in peg parsers." (2016).}
    \bibitem{callbackhell}{Callback Hell, [\url{http://callbackhell.com/}, Accessed on 2020/03/12]}
    \bibitem{material-design}{Material Design Introduction, [\url{https://material.io/design/}, Accessed on 2020/04/09]}
    \bibitem{crafting-interpreters}{Crafting Interpreters, Bob Nystrom, [\url{http://www.craftinginterpreters.com/}, Accessed on 2020/04/09]}
\end{thebibliography}
